The following functions in the header <memory> operate on uninitialized memory.

Copy a range of objects to an unitialized area of memory.

\begin{cpp}
FwdIt uninitialized_copy(InIt first, InIt last, FwdIt out)
FwdIt uninitialized_copy(ExePol pol, FwdIt first, FwdIt last, FwdIt out)
\end{cpp}

Copy count objects to an unitialized area of memory.

\begin{cpp}
FwdIt uninitialized_copy_n(InIt first, Size count, FwdIt out)
FwdIt uninitialized_copy_n(ExePol pol, FwdIt first, Size count, FwdIt out)
\end{cpp}

Copy an object value to an unitialized area of memory.

\begin{cpp}
void uninitialized_fill(FwdIt first, FwdIt last, const T& value)
void uninitialized_fill(ExePol pol, FwdIt first, FwdIt last, const T& value)
\end{cpp}

Copy count object to an unitialized area of memory, starting with first.

\begin{cpp}
void uninitialized_fill_n(FwdIt first, Size count, const T& value)
FwdIt uninitialized_fill_n(FwdIt first, Size count, const T& value)
FwdIt uninitialized_fill_n(ExePol pol, FwdIt first, Size count, const T& value)
\end{cpp}

Move a range of objects to an unitialized area of memory.

\begin{cpp}
FwdIt uninitialized_move(InIt first, InIt last, FwItout)
FwdIt uninitialized_move(ExePol pol, FwdIt first, FwdIt last, FwIt out)
\end{cpp}

Move count objects to an unitialized area of memory.

\begin{cpp}
std::pair<InIt, FwdIt> uninitialized_move_n(InIt first, Size count, FwIt out)
std::pair<FwdIt, FwdIt> uninitialized_move_n(ExePol pol, FwdIt first, Size count,
											 FwIt out)
\end{cpp}

Construct objects by \href{https://en.cppreference.com/w/cpp/language/default_initialization}{default initialization} an unitialized area of memory.

\begin{cpp}
void uninitialized_default_construct(FwdIt first, FwdIt last)
void uninitialized_default_construct(ExePol pol, FwdIt first, FwdIt last)
\end{cpp}

Construct nobjects by \href{https://en.cppreference.com/w/cpp/language/default_initialization}{default initialization} an unitialized area of memory.

\begin{cpp}
FwdIt uninitialized_default_construct_n(FwdIt first, Size n)
FwdIt uninitialized_default_construct_n(ExePol pol, FwdIt first, Size n)
\end{cpp}

Construct objects by \href{https://en.cppreference.com/w/cpp/language/value_initialization}{value initialization} an unitialized area of memory.

\begin{cpp}
void uninitialized_value_construct(FwdIt first, FwdIt last)
void uninitialized_value_construct(ExePol pol, FwdIt first, FwdIt last)
\end{cpp}

Construct n objects by \href{https://en.cppreference.com/w/cpp/language/value_initialization}{value initialization} an unitialized area of memory.

\begin{cpp}
FwdIt uninitialized_value_construct_n(FwdIt first, Size n)
FwdIt uninitialized_value_construct_n(ExePol pol, FwdIt first, Size n)
\end{cpp}

Destroy objects.

\begin{cpp}
constexpr void destroy(FwdIt first, FwdIt last)
void destroy(ExePol pol, FwdIt first, FwdIt last)
\end{cpp}

Destroy n objects.

\begin{cpp}
constexpr void destroy_n(FwdIt first, Size n)
void destroy_n(ExePol pol, FwdIt first, Size n)
\end{cpp}

Destroy an object at a given address p.

\begin{cpp}
constexpr void destroy_at(T* p)
\end{cpp}

Create an object at a given address p.

\begin{cpp}
constexpr T* construct_at(T* p, Args&& ... args)
\end{cpp}












