
std::regex\_replace replaces sequences in a text matching a text pattern. It returns in the simple form std::regex\_replace(text, regex, replString) its result as string. The function replaces an occurrence of regex in text with replString.

\filename{std::replace}

\begin{cpp}
// replace.cpp
...
#include <regex>
...
using namespace std;

string future{"Future"};
string unofficialName{"The unofficial name of the new C++ standard is C++0x."};

regex rgxCpp{R"(C\+\+0x)"};
string newCppName{"C++11"};
string newName{regex_replace(unofficialName, rgxCpp, newCppName)};

regex rgxOff{"unofficial"};
string makeOfficial{"official"};
string officialName{regex_replace(newName, rgxOff, makeOfficial)};

cout << officialName << endl;
							// The official name of the new C++ standard is C++11.
\end{cpp}

In addition to the simple version, C++ has a version of std::regex\_replace working on ranges. It enables you to push the modified string directly into another string:

\begin{cpp}
typedef basic_regex<char> regex;
std::string str2;
std::regex_replace(std::back_inserter(str2),
				   text.begin(), text.end(), regex,replString);
\end{cpp}

All variants of std::regex\_replace have an additional optional parameter. If you set the parameter to std::regex\_constants::format\_no\_copy, you will get the part of the text matching the regular expression. The unmatched text is not copied. If you set the parameter to std::regex\_constants::format\_first\_only std::regex\_replace will only be applied once.



































